---
title: "S1/S2 Phenology Seasonality"
author: "J. Bruce Bell"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/brbell01/Documents/git/atlantic_forest_phenology/R/")
```

Imports
```{r include=FALSE}
# List of libraries
libs <- c("knitr", "tidyr", "lubridate", "clock", "stringr", "dplyr", "ggplot2", "scales", "hrbrthemes", "purrr", "data.table", "tibble", "sf", "tidyquant", "circular", "readr", "DescTools", "padr", "xts", "zoo", "Kendall")

# Loop through the vector and install and load each library if needed
for (lib in libs) {
  if (!require(lib, character.only = TRUE)) {
    install.packages(lib)
    library(lib, character.only = TRUE)
  }
}
```

Setup
```{r}
#set working directory
wd <- getwd()

# Choose indices and sites
S1_band <-"RVI" 
S2_band <- "EVI2"
site_1 <- "rio_doce" ; site_2 <- "sooretama" # site_1 <- "mata_escura" ; site_2 <- "monte_pascoal" # 
site.list <- c(site_1, site_2)

#other params
studyarea_bases <- c("mata_escura","monte_pascoal","rio_doce","sooretama")
S1.band.list <- c("RVI", "VH", "VV")
S2.band.list <- c("EVI2", "EVI", "NDVI")
```

(First run - generate seasonality table) Simple Harmonic Fit (Sentinel-1)
```{r}
# Read in seasonality data from harmonic analysis (S1)
S1_path <- paste0(wd, "/raw/s1/1ha/harmonic/")
S1_files <- dir(S1_path, full.names = T, pattern = "harmonic.csv")
S1_season <- S1_files %>% map(read.csv, header=T)
#loop through and extract seasonality metrics for each file
for (file in 1:length(S1_season)) {
  #fitted
  fit_table <- S1_season[[file]][-c(2)]  # extract table from the file, drop unnecessary columns
  names(fit_table)[1] <- "Date"    # rename date
  fit_table <- fit_table %>% distinct() %>% mutate(Date = date_parse(Date, format = "%b %d, %Y")) # remove duplicates, reformat date
  rownames(fit_table) <- fit_table$Date # format for xts class
  fits_xts <- fit_table %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(fit_table)), "%Y-%m-%d") # xts convert
  
  #trend
  y <- fit_table$fitted ; x <- fit_table$Date ; y <- predict(lm(y~x)) # extract trend
  trend  <- data.frame(cbind(x,y)) ;  colnames(trend) <- c("Date","trend") ; trend$Date <- as.Date(trend$Date) # format trend
  write.csv(trend, paste0(S1_files[[file]] %>% substr(1,nchar(.)-13),"_trend.csv")) # save trend
  # xts convert
  rownames(trend) <- trend$Date ; trend_xts  <- trend %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(trend)), "%Y-%m-%d") 
  
  #detrend
  #fit_detrend <- fit_table %>% diff() %>% fortify() ; colnames(fit_detrend) <- c("Date","detrend")
  #write.csv(fit_detrend, paste0(S1_files[[file]] %>% substr(1,nchar(.)-13),"_detrend.csv")) # save detrended file
  
  # use detrended hereafter for seasonality definition ; add trend values to final table
  detrend_xts <- fits_xts - trend_xts  ; detrend <- detrend_xts %>% fortify() ; colnames(detrend) <- c("Date","detrended")
  jan1 <- detrend$Date %>% year() %>% unique()  # list of years
  date_breaks <- paste0(jan1, "-01-01") %>% as.Date()  # Jan. 1 of each year
  seasonality <- data.frame(matrix(NA, nrow=1, ncol=3))  # format a blank seasonality table to hold metrics
  dates <- matrix(rep("2017-01-01"),3) %>% t %>% data.frame() %>% mutate_all(as.Date)
  seasonality <- seasonality %>% cbind(dates)  # append date columns
  names(seasonality) <- c("season_start_val", "peak_val", "season_end_val", "season_start_date", "peak_date", "season_end_date") # col names
  seasonality <- seasonality[-1,] # remove dummy values

#define function which limits the dataframe to a single calendar year
    single_year_filter <- function (dt)  {
                dt %>% pad() %>% thicken(paste0(12, " days")) %>% select(-Date) %>% rename_at(ncol(.), ~"Date") %>% 
                       mutate(Date = Date+1) %>% relocate(Date) %>% pad() %>%
                       filter(between(Date, date_breaks[year], if_else(!is.na(date_breaks[year+1]), date_breaks[year+1], tail(fit_table$Date, n=1))))
                 }
  #loop through each year and get season peak, start, end dates from the detrended series, and values from the original time series 
  for (year in 1:length(date_breaks)) {
      dtable <- detrend %>% single_year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(detrended)) %>% distinct(Date, .keep_all = TRUE)
      ttable <- trend %>% single_year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(trend)) %>% distinct(Date, .keep_all = TRUE)
      otable <- fit_table %>% single_year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(fitted)) %>% distinct(Date, .keep_all = TRUE)
      min_detr <- min(dtable$detrended, na.rm=T) # get season minimum detrended value
      min_date <- dtable %>% filter(detrended == min_detr) # get season minimum date
      min_date <- first(min_date[[1]])  # enforce single date
      min <- otable %>% filter(Date == min_date) ; min <- min$fitted # get season minimum original fitted value
      min_tr <- ttable %>% filter(Date == min_date) ; min_tr <- min_tr$trend # get trend value corresp to min. date
      max_detr <- max(dtable$detrended, na.rm=T) # get season peak value
      max_date <- dtable %>% filter(detrended == max_detr)  # get season peak date
      max_date <- first(max_date[[1]]) # enforce single date
      max <- otable %>% filter(Date == max_date) ; max <- max$fitted # get season peak original fitted value
      max_tr <- ttable %>% filter(Date == max_date) ; max_tr <- max_tr$trend # get trend value corresp to min. date
       # narrow season start window
      if (min_date < max_date) {min_max <- otable %>% filter(between(Date, min_date, max_date))
      } else {min_max <- otable %>% filter(between(Date, min_date, max_date + 365/2))}
      mean <- Closest(min_max$fitted, (max_tr + min_tr)/2, na.rm = T)[1] # find season start value, enforce single value
      s_start <- min_max %>% filter(fitted == mean) # get season start date
      s_start <- ifelse(!is.na(mean), s_start$Date, NA) %>% as.Date() #get season start date with case for final year
      #get season end date with cases for late peaks, early peaks, and final yr
      if (!is.na(mean) && max_date > s_start) {s_end <- max_date - s_start + max_date %>% as.Date()
        } else if (!is.na(mean) && max_date < s_start) {s_end <- s_start - max_date + min_date %>% as.Date() 
        } else {s_end <- NA}
      # find season end value, with final year case
      if (!is.na(mean)) {
        s_end_val <- fit_table %>% filter(between(Date, s_end-6, s_end+6))
        } else {s_end <- NA}
      if (!is.na(mean)) {  
        s_end_val <- s_end_val$fitted %>% mean # enforce single value
      } else {s_end_val <- NA}  
      seasonality[nrow(seasonality) + 1,] <- list(mean, max, s_end_val, s_start, max_date, s_end)
  }
  
  #lookup tables for original values (not the detrend values used above)
  #fit_original <- fit_table %>% fortify() ; colnames(fit_table) <- c("Date","fitted")
  #fit_original <- fit_original[,c("Date","fitted")]
   
  #pivot table for lookup 
  #vals_long <- seasonality[,1:3] %>% pivot_longer(everything(), names_to = "val_type", values_to = "value")
  #dates <- seasonality[,4:6]
  #dates_long <- dates %>% pivot_longer(everything(), names_to = "date_type", values_to = "Date")
  #join for lookup
 # season_new <- cbind(vals_long,dates_long) %>% left_join(fit_original, by="Date")
  #undo pivot
 # new_vals <- season_new %>% select(c(val_type, fitted_original)) %>% mutate(grp = rep(1:3, each=6))
 # new_vals <- cbind(grp = unique(new_vals$grp), unstack(new_vals, fitted_original ~ val_type)) %>% select(-grp)
  
  #gather new values
 # seasonality <- cbind(dates, new_vals)
  
  # Add variables to seasonality table 
  seasonality <- seasonality %>% mutate("Seas." = seq(1:length(season_start_val)))
  seasonality <- seasonality %>% mutate("Ampl." = max-min)
  seasonality <- seasonality %>% mutate("length_days" = ymd(season_end_date)-ymd(season_start_date))
  
  # Season definition 1: Peak to Jan. 1
  seasonality <- seasonality %>% mutate("phase_peak" = 
                             ifelse(year(peak_date) == year(season_start_date),
                             ymd(peak_date) - floor_date(ymd(peak_date), unit="year"),
                             ymd(peak_date) - floor_date(ymd(peak_date), unit="year") + 365))
  
  # write out seasonality file to disk
  study_site_names <- rep(c("mata_escura","monte_pascoal","rio_doce","sooretama"), each=length(S1.band.list))
  band_list <- rep(S1.band.list, length(study_site_names))
  write.csv(seasonality, paste0(S1_path, study_site_names[[file]],"_s1_1ha_", band_list[[file]], "_seasonality_processed.csv"))
}
```

(First run - generate seasonality table) Simple Harmonic Fit (Sentinel-2)
```{r}
start_date <- as.Date("2019-01-01")  # start analysis from this date
# Read in seasonality data from harmonic analysis (S2)
S2_path <- paste0(wd, "/raw/s2/1ha/harmonic/")
S2_files <- dir(S2_path, full.names = T, pattern = "harmonic.csv")
S2_season <- S2_files %>% map(read.csv, header=T)

#define year filter function
    year_filter <- function (dt)  {
                dt %>% pad() %>% thicken(paste0(5, " days")) %>% select(-Date) %>% rename_at(ncol(.), ~"Date") %>% 
                       mutate(Date = Date+1) %>% relocate(Date) %>% pad() %>% 
                       filter(between(Date, date_breaks[year], if_else(!is.na(date_breaks[year+1]), date_breaks[year+1], tail(fit_table$Date, n=1))))
                 }
#loop through and extract seasonality metrics for each file
for (file in 1:length(S2_season)) {
  #fitted
  fit_table <- S2_season[[file]][-c(2)]  # extract table from the file, drop unnecessary columns
  names(fit_table)[1] <- "Date"    # rename date
  fit_table <- fit_table %>% distinct() %>% mutate(Date = date_parse(Date, format = "%b %d, %Y")) # remove duplicates, reformat date
  fit_table <- fit_table %>% filter(Date >= start_date)
  rownames(fit_table) <- fit_table$Date # format for xts class
  fits_xts <- fit_table %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(fit_table)), "%Y-%m-%d") # xts convert
  
  #trend
  y <- fit_table$fitted ; x <- fit_table$Date ; y <- predict(lm(y~x)) # extract trend
  trend  <- data.frame(cbind(x,y)) ;  colnames(trend) <- c("Date","trend") ; trend$Date <- as.Date(trend$Date) # format trend
#  write.csv(trend, paste0(S2_files[[file]] %>% substr(1,nchar(.)-13),"_trend.csv")) # save trend
  rownames(trend) <- trend$Date ; trend_xts  <- trend %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(trend)), "%Y-%m-%d") # xts convert
  
  #detrend
  #fit_detrend <- fit_table %>% diff() %>% fortify() ; colnames(fit_detrend) <- c("Date","detrend")
  #write.csv(fit_detrend, paste0(S2_files[[file]] %>% substr(1,nchar(.)-13),"_detrend.csv")) # save detrended file
  
  # use detrended hereafter for seasonality definition ; add trend values to final table
  detrend_xts <- fits_xts - trend_xts  ; detrend <- detrend_xts %>% fortify() ; colnames(detrend) <- c("Date","detrended")
  jan1 <- detrend$Date %>% year() %>% unique()  # list of years
  date_breaks <- paste0(jan1, "-01-01") %>% as.Date()  # Jan. 1 of each year
  seasonality <- data.frame(matrix(NA, nrow=1, ncol=3))  # format a blank seasonality table to hold metrics
  dates <- matrix(rep("2017-01-01"),3) %>% t %>% data.frame() %>% mutate_all(as.Date)
  seasonality <- seasonality %>% cbind(dates)  # append date columns
  names(seasonality) <- c("season_start_val", "peak_val", "season_end_val", "season_start_date", "peak_date", "season_end_date") # col names
  seasonality <- seasonality[-1,] # remove dummy values

    #loop through each year and get season peak, start, end dates from the detrended series, and values from the original time series 
  for (year in 1:length(date_breaks)) {
      dtable <- detrend %>% year_filter() %>% group_by(Date) 
                %>% arrange(.by_group = TRUE, desc(detrended)) %>% distinct(Date, .keep_all = TRUE)
      ttable <- trend %>% year_filter() %>% group_by(Date) %>% 
                arrange(.by_group = TRUE, desc(trend)) %>% distinct(Date, .keep_all = TRUE)
      otable <- fit_table %>% year_filter() %>% group_by(Date) %>% 
                arrange(.by_group = TRUE, desc(fitted)) %>% distinct(Date, .keep_all = TRUE)
      min_detr <- min(dtable$detrended, na.rm=T) # get season minimum detrended value
      min_date <- dtable %>% filter(detrended == min_detr) # get season minimum date
      min_date <- first(min_date[[1]])  # enforce single date
      min <- otable %>% filter(Date == min_date) ; min <- min$fitted # get season minimum original fitted value
      min_tr <- ttable %>% filter(Date == min_date) ; min_tr <- min_tr$trend # get trend value corresp to min. date
      max_detr <- max(dtable$detrended, na.rm=T) # get season peak value
      max_date <- dtable %>% filter(detrended == max_detr)  # get season peak date
      max_date <- first(max_date[[1]]) # enforce single date
      max <- otable %>% filter(Date == max_date) ; max <- max$fitted # get season peak original fitted value
      max_tr <- ttable %>% filter(Date == max_date) ; max_tr <- max_tr$trend # get trend value corresp to min. date
       # narrow season start window
      if (min_date < max_date) {min_max <- otable %>% filter(between(Date, min_date, max_date))
      } else {min_max <- otable %>% filter(between(Date, min_date, max_date + 365))}
      mean <- Closest(min_max$fitted, (max_tr + min_tr)/2, na.rm = T)[1] # find season start value, enforce single value
      # get season start date
      s_start <- min_max %>% filter(fitted == mean) 
      #get season start date with case for final year
      s_start <- ifelse(!is.na(mean), s_start$Date, NA) %>% as.Date()
      #get season end date with cases for late peaks, early peaks, and final yr
      if (!is.na(mean) && max_date > s_start) {s_end <- max_date - s_start + max_date %>% as.Date()
        } else if (!is.na(mean) && max_date < s_start) {s_end <- s_start - max_date + min_date %>% as.Date() 
        } else {s_end <- NA}
      # find season end value, with final year case
      if (!is.na(mean)) {
        s_end_val <- fit_table %>% filter(between(Date, s_end-3, s_end+3))
        } else {s_end <- NA}
      if (!is.na(mean)) {  
        s_end_val <- s_end_val$fitted %>% mean # enforce single value
      } else {s_end_val <- NA}  
      seasonality[nrow(seasonality) + 1,] <- list(mean, max, s_end_val, s_start, max_date, s_end)
  }
  
  # Add variables to seasonality table 
  seasonality <- seasonality %>% mutate("Seas." = seq(1:length(season_start_val)))
  seasonality <- seasonality %>% mutate("Ampl." = max-min)
  seasonality <- seasonality %>% mutate("length_days" = ymd(season_end_date)-ymd(season_start_date))
  
  # Season definition 1: Peak to Jan. 1
  seasonality <- seasonality %>% mutate("phase_peak" = 
                             ifelse(year(peak_date) == year(season_start_date),
                             ymd(peak_date) - floor_date(ymd(peak_date), unit="year"),
                             ymd(peak_date) - floor_date(ymd(peak_date), unit="year") + 365))

  # write out seasonality file to disk
  study_site_names <- rep(c("mata_escura","monte_pascoal","rio_doce","sooretama"), each=length(S2.band.list))
  band_list <- rep(S2.band.list, length(study_site_names))
  write.csv(seasonality, paste0(S2_path, study_site_names[[file]],"_s2_1ha_", band_list[[file]], "_seasonality_processed.csv"))
}
```

Quick plot to show seasonality metrics reconstructed from harmonic table above
```{r}
  S1_path <- paste0(wd, "/raw/s1/1ha/harmonic/")
  S1_files <- dir(S1_path, full.names = T, pattern = "harmonic.csv")
  fit_table <- S1_files[1] %>% read.csv(header=T)
  names(fit_table)[1] <- "Date"    # rename date
  fit_table <- fit_table[-c(2)]  # extract table from the file, drop unnecessary columns
  fit_table <- fit_table %>% distinct() %>% mutate(Date = date_parse(Date, format = "%b %d, %Y"))
  rownames(fit_table) <- fit_table$Date # format for xts class
  fits_xts <- fit_table %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(fit_table)), "%Y-%m-%d") # xts convert
  
  #trend
  y <- fit_table$fitted ; x <- fit_table$Date ; y <- predict(lm(y~x)) # extract trend
  trend  <- data.frame(cbind(x,y)) ;  colnames(trend) <- c("Date","trend") ; trend$Date <- as.Date(trend$Date) # format trend
  rownames(trend) <- trend$Date ; trend_xts  <- trend %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(trend)), "%Y-%m-%d") # xts convert
  
  # use detrended hereafter for seasonality definition ; add trend values to final table
  detrend_xts <- fits_xts - trend_xts  ; detrend <- detrend_xts %>% fortify() ; colnames(detrend) <- c("Date","detrended")
  jan1 <- detrend$Date %>% year() %>% unique()  # list of years
  date_breaks <- paste0(jan1, "-01-01") %>% as.Date()  # Jan. 1 of each year
  seasonality <- data.frame(matrix(NA, nrow=1, ncol=3))  # format a blank seasonality table to hold metrics
  dates <- matrix(rep("2017-01-01"),3) %>% t %>% data.frame() %>% mutate_all(as.Date)
  seasonality <- seasonality %>% cbind(dates)  # append date columns
  names(seasonality) <- c("season_start_val", "peak_val", "season_end_val", "season_start_date", "peak_date", "season_end_date") # col names
  seasonality <- seasonality[-1,] # remove dummy values
 
  year_filter <- function (dt)  {
                dt %>% pad() %>% thicken(paste0(12, " days")) %>% select(-Date) %>% rename_at(ncol(.), ~"Date") %>% 
                       mutate(Date = Date+1) %>% relocate(Date) %>% pad() %>% 
                       filter(between(Date, date_breaks[year], if_else(!is.na(date_breaks[year+1]), date_breaks[year+1], tail(fit_table$Date, n=1))))
                 }
  #loop through each year and get season peak, start, end dates from the detrended series, and values from the original time series 
  for (year in 1:length(date_breaks)) {
      dtable <- detrend %>% year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(detrended)) %>% distinct(Date, .keep_all = TRUE)
      ttable <- trend %>% year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(trend)) %>% distinct(Date, .keep_all = TRUE)
      otable <- fit_table %>% year_filter() %>% group_by(Date) %>% arrange(.by_group = TRUE, desc(fitted)) %>% distinct(Date, .keep_all = TRUE)
      min_detr <- min(dtable$detrended, na.rm=T) # get season minimum detrended value
      min_date <- dtable %>% filter(detrended == min_detr) # get season minimum date
      min_date <- first(min_date[[1]])  # enforce single date
      min <- otable %>% filter(Date == min_date) ; min <- min$fitted # get season minimum original fitted value
      min_tr <- ttable %>% filter(Date == min_date) ; min_tr <- min_tr$trend # get trend value corresp to min. date
      max_detr <- max(dtable$detrended, na.rm=T) # get season peak value
      max_date <- dtable %>% filter(detrended == max_detr)  # get season peak date
      max_date <- first(max_date[[1]]) # enforce single date
      max <- otable %>% filter(Date == max_date) ; max <- max$fitted # get season peak original fitted value
      max_tr <- ttable %>% filter(Date == max_date) ; max_tr <- max_tr$trend # get trend value corresp to min. date
       # narrow season start window
      if (min_date < max_date) {min_max <- otable %>% filter(between(Date, min_date, max_date))
      } else {min_max <- otable %>% filter(between(Date, max_date, min_date))}
      mean <- Closest(min_max$fitted, (max_tr + min_tr)/2, na.rm = T)[1] # find season start value, enforce single value
      s_start <- min_max %>% filter(fitted == mean)  # get season start date
      s_start <- ifelse(!is.na(mean), s_start$Date, NA) %>% as.Date() #get season start date with case for final year
      #get season end date with case for final yr
      if (!is.na(mean) && max_date > s_start) {s_end <- max_date - s_start + max_date %>% as.Date()
        } else if (!is.na(mean) && max_date < s_start) {s_end <- s_start - max_date + min_date %>% as.Date() 
        } else {s_end <- NA}
      # find season end value, with final year case
      if (!is.na(mean)) {
        s_end_val <- fit_table %>% filter(between(Date, s_end-6, s_end+6))
        } else {s_end <- NA}
      if (!is.na(mean)) {  
        s_end_val <- s_end_val$fitted %>% mean # enforce single value
      } else {s_end_val <- NA}  
      seasonality[nrow(seasonality) + 1,] <- list(mean, max, s_end_val, s_start, max_date, s_end)
  }
  
  #pivot table for lookup 
  vals <- seasonality[,1:3] %>% pivot_longer(everything(), names_to = "val_type", values_to = "value")
  dates <- seasonality[,4:6] %>% pivot_longer(everything(), names_to = "date_type", values_to = "Date")
  seasonality_new <- cbind(vals,dates) ; rownames(seasonality_new) <- seasonality_new$Date # format for xts class
  seasonality_xts <- seasonality_new %>% select(-c(Date)) %>% select("value") %>% xts(order.by = as.Date(rownames(seasonality_new)), "%Y-%m-%d")
  

  #plot
  fits_xts %>% plot(col="blue")                # original fitted values
  points(fits_xts, col="blue", cex = 1)
  points(seasonality_xts, col="red", cex = 2)    # calculated seasonality metrics
  lines(trend_xts, lty=2, lwd=2.5, col="red")

    
  #save to file
  png(file="seasonality_metrics_methodology.png", width=1000, height=600)
  fits_xts %>% plot(col="blue")                # original fitted values
  points(fits_xts, col="blue", cex = 1)
  points(seasonality_xts, col="red", cex = 2)    # calculated seasonality metrics
  lines(trend_xts, lty=2, lwd=2.5, col="red")
  dev.off()
```

Read in seasonality data. Run A (single harmonic modeled) or B (multiple harmonic modeled)

A. Read in fitted and trend for time series S1/S2 plotting (single harmonic)
```{r}
hnumber <- "single"
S1_start_date <- as.Date("2016-01-10")
S2_start_date <- as.Date("2018-12-17")
  
# Read fitted time series and trend data files and append
S1_dir <- "/raw/s1/1ha/harmonic/"
S2_dir <- "/raw/s2/1ha/harmonic/"

S1_fitted <- dir(paste0(wd, S1_dir), full.names = T, pattern = paste0(S1_band,"_2016_2022_harmonic.csv")) %>% map(read.csv, row.names = NULL)
S1_trend <- dir(paste0(wd, S1_dir), full.names = T, pattern = paste0(S1_band,"_2016_2022_trend.csv")) %>% map(read.csv, row.names = NULL)
S2_fitted <- dir(paste0(wd, S2_dir), full.names = T, pattern = paste0(S2_band,"_2016_2022_harmonic.csv")) %>% map(read.csv, row.names = NULL)
S2_trend <- dir(paste0(wd, S2_dir), full.names = T, pattern = paste0(S2_band,"_2016_2022_trend.csv")) %>% map(read.csv, row.names = NULL)

fittedlist <- list() ; trendlist <- list()

#loop through each of the sites and grab the veg. index data
for (site in site.list) {
  #sentinel-1
  S1_fitteddata <- vector(mode = "list", length = 0)
  S1_table <- S1_fitted[[match(site,studyarea_bases)]]
  S1_site_band <- paste0(site, "_", S1_band)
  S1_fitteddata[[S1_site_band]] <- S1_table 
  names(S1_fitteddata[[S1_site_band]]) <- c("Date", S1_band, "fitted")
  S1_fitteddata[[S1_site_band]] <- S1_fitteddata[[S1_site_band]] %>% select(-all_of(S1_band)) %>%
  mutate(Date = date_parse(Date, format = "%b %d, %Y")) %>% 
  group_by(Date) %>% summarize(fitted = mean(fitted, na.rm = TRUE)) %>% ungroup()
  
  #sentinel-1 trend
  S1_trenddata <- vector(mode = "list", length = 0)
  S1_trendtable <- S1_trend[[match(site,studyarea_bases)]] %>% select(-X)
  S1_trendtable <- data.frame(cbind(S1_trendtable$Date, S1_trendtable$trend)) # strange behavior but this is necessary for some reason
  S1_trenddata[[S1_site_band]] <- S1_trendtable
  names(S1_trenddata[[S1_site_band]]) <- c("Date", "trend")
  S1_trenddata[[S1_site_band]] <- S1_trenddata[[S1_site_band]] %>% mutate(trend = as.numeric(trend))
  S1_trenddata[[S1_site_band]] <- S1_trenddata[[S1_site_band]] %>% mutate(Date = as.Date(Date)) %>% 
                                  group_by(Date) %>% summarize(trend = mean(trend, na.rm = TRUE)) %>% ungroup()

  #sentinel-2
  S2_fitteddata <- vector(mode = "list", length = 0)
  S2_table <- S2_fitted[[match(site,studyarea_bases)]]
  S2_site_band <- paste0(site, "_", S2_band)
  S2_fitteddata[[S2_site_band]] <- S2_table 
  names(S2_fitteddata[[S2_site_band]]) <- c("Date", S2_band, "fitted")
  S2_fitteddata[[S2_site_band]] <- S2_fitteddata[[S2_site_band]] %>% select(-all_of(S2_band)) %>%
  mutate(Date = date_parse(Date, format = "%b %d, %Y")) %>% group_by(Date) %>% summarize(fitted = mean(fitted, na.rm = TRUE)) %>% ungroup()

  #sentinel-2 trend
  S2_trenddata <- vector(mode = "list", length = 0)
  S2_trendtable <- S2_trend[[match(site,studyarea_bases)]] %>% select(-X)
  S2_trendtable <- data.frame(cbind(S2_trendtable$Date, S2_trendtable$trend)) # strange behavior but this is necessary for some reason
  S2_trenddata[[S2_site_band]] <- S2_trendtable
  names(S2_trenddata[[S2_site_band]]) <- c("Date", "trend")
  S2_trenddata[[S2_site_band]] <- S2_trenddata[[S2_site_band]] %>% mutate(trend = as.numeric(trend))
  S2_trenddata[[S2_site_band]] <- S2_trenddata[[S2_site_band]] %>% mutate(Date = as.Date(Date)) %>% 
                                  group_by(Date) %>% summarize(trend = mean(trend, na.rm = TRUE)) %>% ungroup()
  
  #combine S1/S2
  fittedlist[[site]] <- list(S1_fitteddata[[S1_site_band]], S2_fitteddata[[S2_site_band]])
  trendlist[[site]] <- list(S1_trenddata[[S1_site_band]], S2_trenddata[[S2_site_band]])
}
# keep a copy for model comparison
fittedlist -> single_harmonic_fittedlist
```

B. Read in fitted and trend Timesat outputs for time series S1/S2 plotting (multiple harmonic)
```{r}
hnumber <- "multiple"
S1_orb82_start_date <- as.Date("2016-01-10")
S1_orb155_start_date <- as.Date("2016-01-03")
S2_start_date <- as.Date("2018-12-17")
  
# Read fitted time series and trend data files and append
season_directory <- "/timesat/output/third_run/"
TS_directory <- "/timesat/formatted_timeseries/third_run/"

S1_fitted <- dir(paste0(wd, season_directory, "S1"), full.names = T, pattern = paste0(S1_band,"_STL_season.txt")) %>% map(fread, sep=" ")
S1_trend <- dir(paste0(wd, season_directory, "S1"), full.names = T, pattern = paste0(S1_band,"_STL_trend.txt")) %>% map(fread, sep=" ")
S2_fitted <- dir(paste0(wd, season_directory, "S2"), full.names = T, pattern = paste0(S2_band,"_STL_season.txt")) %>% map(fread, sep=" ")
S2_trend <- dir(paste0(wd, season_directory, "S2"), full.names = T, pattern = paste0(S2_band,"_STL_trend.txt")) %>% map(fread, sep=" ")

fittedlist <- list() ; trendlist <- list()

#loop through both of the sites and grab the veg. index data
for (site in site.list) {
  #sentinel-1 fitted
  S1_fitteddata <- vector(mode = "list", length = 0)
  S1_table <- S1_fitted[[match(site,studyarea_bases)]] %>% t() %>% data.frame()
  colnames(S1_table) <- "fitted"
  len <- S1_table[,1] %>% length()
  if(site == "rio_doce") {
    dates <- seq(S1_orb155_start_date, by = "12 days", length.out=len)
  } else {
    dates <- seq(S1_orb82_start_date, by = "12 days", length.out=len)}
  S1_table <- S1_table %>% mutate(Date = dates) %>% relocate(Date)
  S1_site_band <- paste0(site, "_", S1_band)
  S1_fitteddata[[S1_site_band]] <- S1_table
  #sentinel-1 trend
  S1_trenddata <- vector(mode = "list", length = 0)
  S1_trendtable <- S1_trend[[match(site,studyarea_bases)]] %>% t() %>% data.frame()
  colnames(S1_trendtable) <- "trend"
  len <- S1_trendtable[,1] %>% length()
  S1_trendtable <- S1_trendtable %>% mutate(Date = dates) %>% relocate(Date)
  S1_trenddata[[S1_site_band]] <- S1_trendtable
  
  #sentinel-2 fitted
  S2_fitteddata <- vector(mode = "list", length = 0)
  S2_table <- S2_fitted[[match(site,studyarea_bases)]] %>% t() %>% data.frame()
  colnames(S2_table) <- "fitted"
  len <- S2_table[,1] %>% length()
  dates <- seq(S2_start_date, by = "5 days", length.out=len)
  S2_table <- S2_table %>% mutate(Date = dates) %>% relocate(Date)
  S2_site_band <- paste0(site, "_", S2_band)
  S2_fitteddata[[S2_site_band]] <- S2_table
  #sentinel-2 trend
  S2_trenddata <- vector(mode = "list", length = 0)
  S2_trendtable <- S2_trend[[match(site,studyarea_bases)]] %>% t() %>% data.frame()
  colnames(S2_trendtable) <- "trend"
  len <- S2_trendtable[,1] %>% length()
  S2_trendtable <- S2_trendtable %>% mutate(Date = dates) %>% relocate(Date)
  S2_trenddata[[S2_site_band]] <- S2_trendtable
  
  #combine S1/S2
  fittedlist[[site]] <- list(S1_fitteddata[[S1_site_band]], S2_fitteddata[[S2_site_band]])
  trendlist[[site]] <- list(S1_trenddata[[S1_site_band]], S2_trenddata[[S2_site_band]])
}

# keep a copy for model comparison
fittedlist -> multiple_harmonic_fittedlist
```

A/B Plot full fitted times series S1/S2 (x = time, y = unscaled indices)
```{r}
#Define study site
study_site <-"sooretama" # "sooretama" # "rio_doce" # "monte_pascoal" # "mata_escura"

if (study_site == "rio_doce") {studyarea <-  "PE Rio Doce, MG"
} else if (study_site == "sooretama") {studyarea <- "REBIO Sooretama, ES"
} else if (study_site == "monte_pascoal") {studyarea <- "PNH Monte Pascoal, BA"
} else if (study_site == "mata_escura") {studyarea <- "REBIO Mata Escura, MG"
} else {studyarea <- "NA"}

if (study_site == "rio_doce" | study_site == "mata_escura") {foresttype <- "seasonal semi-deciduous forest"
} else if (study_site == "monte_pascoal" | study_site == "sooretama") {foresttype <- "broadleaf evergreen forest"
} else {foresttype <- "NA"}

S1fit <- fittedlist[[study_site]][[1]] %>% 
            select(Date, colnames(fittedlist[[study_site]][[1]]) %>% str_which(.,"fitted") %>% first()) %>% distinct()
S2fit <- fittedlist[[study_site]][[2]] %>% 
            select(Date, colnames(fittedlist[[study_site]][[2]]) %>% str_which(.,"fitted") %>% last()) %>% distinct()

data <- full_join(S1fit, S2fit, by="Date")

xlims <- c(as.Date("2019-01-01"), as.Date("2021-12-31"))
month_breaks <- date_seq(from = floor_date(xlims[1], unit="month"), to = floor_date(xlims[2], unit="month"), by = duration_months(1))
mon_abbr <- month_breaks %>% format("%b") %>% substr(1,1)
date_frame <- data.frame("Date" = month_breaks, mon_abbr) %>% as_tibble()

data <- full_join(data, date_frame, by="Date") %>% arrange(Date)

colnames(data) <- c("Date", "RVI_fitted", "EVI2_fitted", "mon_abbr")
jan1 <- paste0(data$Date %>% year %>% unique, "-01-01") %>% as.Date()

meanRVI <- mean(data$RVI_fitted, na.rm = T)
meanEVI2 <- mean(data$EVI2_fitted, na.rm=T)
data$EVI2_fitted_rescaled <- (data$EVI2_fitted - meanEVI2)/2 + meanRVI
data <- data %>% select(-EVI2_fitted)
long_data <- pivot_longer(data, cols = c(-Date, -mon_abbr), names_to = "Index_Type", values_to = "Value", names_prefix = "")
long_data$Index_Type <- factor(long_data$Index_Type)

S1legend <- paste0("all plots mean ", S1_band) ; S2legend <- paste0("all plots mean ", S2_band)
colors <- c("all plots mean RVI" = "blue", "all plots mean EVI2" = "darkgreen")
#linetypes <- c("all plots mean RVI" = "dashed", "all plots mean EVI2" = "dashed")

colors <- c("darkgreen", "blue")


#plot time series
ggplot(data = long_data, aes(x=Date)) +  
  geom_ma(size = 3, ma_fun = SMA, n = 12, aes(y = Value, color = Index_Type)) +
 # geom_line(aes(y= EVI2_fitted+0.2859671, colour = "all plots mean EVI2"), linetypes = "dashed", na.rm = T) +
 # geom_ma(ma_fun = SMA, n = 12, aes(y= RVI_fitted, colour = "all plots mean RVI"), linetypes = 1) +
 # geom_ma(ma_fun = SMA, n = 10, aes(y= EVI2_fitted, colour = "all plots mean EVI2"), linetypes = 2) +
  scale_color_manual(name = "", values= colors, labels = c(S2legend, S1legend)) +
  theme_pubr() +
#  scale_linetype_manual(name = "", values= "dashed") +
#       theme_minimal(base_size = 14) +
       scale_x_date(date_labels = mon_abbr
               , limits = xlims
               , date_breaks = "1 month"
#               , minor_breaks = "1 day"
               , expand = expansion(add = .05) 
                ) +
       scale_y_continuous(breaks = pretty_breaks()
              , name = "Sentinel-1 RVI Index" 
              , sec.axis = sec_axis(~((.-meanRVI)*2)+meanEVI2, name="Sentinel-2 EVI2 Index", breaks = pretty_breaks())) +
       xlab("Month, Year") +
       theme(legend.position = "top"
             , axis.text.x = element_text(angle = 0)
             , axis.line.y.left = element_line(color = "blue")
             , axis.line.y.right = element_line(color = "darkgreen")
             , axis.text.y = element_text(color = "blue")
             , axis.title.y = element_text(color = "blue")
             , axis.text.y.right = element_text(color = "darkgreen")
             , axis.title.y.right = element_text(color = "darkgreen")) +
       geom_vline(color = "red", alpha = 0.3, xintercept = jan1) +
       geom_text(aes(x = as.Date(paste0(format(Date, "%Y"), "-01-01")), y = -Inf, label = format(Date, "%Y")), angle = 90, hjust = 0, vjust = 0, color = "black", size = 4) +
      # geom_text(aes(y = 1, x=ymd(jan1[[1]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
      # geom_text(aes(y = 1, x=ymd(jan1[[2]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
      # geom_text(aes(y = 1, x=ymd(jan1[[3]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
       ggtitle(paste0("Sentinel-1 & Sentinel-2 modeled seasonality over \n", studyarea, " (single harmonic)"))
```

Save plot
```{r}
ggsave(paste0(study_site, "_fitted_seasonality_", S1_band, "_", S2_band,"_", hnumber, "_harmonic.png"), units="in", width=10, height=5, dpi=300, device = 'png')
```

A/B Plot full fitted times series S1/S2 (x = time, y = rescaled indices)
```{r}
#Define study site
site <-"rio_doce" # "sooretama" # "rio_doce" # "monte_pascoal" # "mata_escura"

if (site == "rio_doce") {studyarea <-  "PE Rio Doce, MG"
} else if (site == "sooretama") {studyarea <- "REBIO Sooretama, ES"
} else if (site == "monte_pascoal") {studyarea <- "PNH Monte Pascoal, BA"
} else if (site == "mata_escura") {studyarea <- "REBIO Mata Escura, MG"
} else {studyarea <- "NA"}

if (site == "rio_doce" | site == "mata_escura") {foresttype <- "seasonal semi-deciduous forest"
} else if (site == "monte_pascoal" | site == "sooretama") {foresttype <- "broadleaf evergreen forest"
} else {foresttype <- "NA"}

S1fit <- fittedlist[[site]][[1]] %>% 
            select(Date, colnames(fittedlist[[site]][[1]]) %>% str_which(.,"fitted") %>% first()) %>% distinct()
S2fit <- fittedlist[[site]][[2]] %>% 
            select(Date, colnames(fittedlist[[site]][[2]]) %>% str_which(.,"fitted") %>% last()) %>% distinct()

#detrend and scale
rownames(S1fit) <- S1fit$Date ; rownames(S2fit) <- S2fit$Date
S1scaled <- S1fit %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S1fit)), "%Y-%m-%d") %>% diff() %>% scale()
S1scale <- scale(S1fit[[2]]) %>% attr("scaled:scale") ; S1center <- scale(S1fit[[2]])  %>% attr("scaled:center")
S2scaled <- S2fit %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S2fit)), "%Y-%m-%d") %>% diff() %>% scale()
S2scale <- scale(S2fit[[2]]) %>% attr("scaled:scale") ; S2center <- scale(S2fit[[2]])  %>% attr("scaled:center")

scaled <- merge(S1scaled, S2scaled)
#  %>% pivot_longer(!Date, names_to = "series") %>% mutate(Date = as.Date(Date))

#data <- data.frame(dates, S1fit, S2fit) 

S1legend <- paste0("all plots mean", S1_band) ; S2legend <- paste0("all plots mean", S2_band)
colors <- c("RVI_fitted" = "blue", "EVI2_fitted" = "darkgreen")
linetypes <- c("RVI_fitted" = "dotted", "EVI2_fitted" = "dashed")

xlims <- c(as.Date("2019-01-01"), as.Date("2022-01-01"))
#ylims <- c(-10,10)

scaled <- fortify(scaled) ; S1_adjust <- 0.8*scaled[,3]
scaled <- scaled %>% mutate(S1_band = S1_adjust) ; scaled <- scaled[-3]
colnames(scaled) <- c("Date", "RVI_fitted", "EVI2_fitted")
jan1 <- paste0(scaled$Date %>% year %>% unique, "-01-01") %>% as.Date()
scaled <- scaled %>% pivot_longer(names(scaled)[-1], names_to = "index")
#read in rainfall data
rainfall_filename <- paste0(str_to_title(site), "_IMERG_2016-2022_14day_sum.csv")  ;

#Import max. hourly rate (reported daily) rainfall data ; rainfall is raw and filtered for values 
raintable <- read.csv(paste0("/Users/brbell01/Google Drive/PHD/CUNY/Research/Data/Raster/IMERG/", rainfall_filename), header=T)

raintable <- raintable %>% mutate(Date = date_parse(raintable$system.time_start, format = "%b %d, %Y")) %>% select(-system.time_start) %>% relocate(Date) %>% mutate(month = month(ymd(Date)), year = year(ymd(Date))) %>% group_by(month, year) %>% summarise(monthly_precip = sum(precipitationCal)) %>% ungroup() %>% mutate(Date = paste0(year,"-",month,"-01"))
raintable$Date <- ymd(raintable$Date)

precip_scaled <- scale(raintable$monthly_precip)[,1]
raintable <-raintable %>% mutate(precip_scaled = precip_scaled)
raintable <-raintable %>% mutate(precip_baseline = -1.5)

#plot time series
ggplot() +  
 # geom_point(data = raintable, aes(x=Date, y=precip_scaled/2-1), shape = c("—"), size = 5) +
  geom_segment(data = raintable, aes(x=Date, xend=Date, y=precip_scaled/2-1, yend=precip_baseline), size=5, alpha=0.4) +
  geom_text(data = raintable, aes(x=Date, y=precip_scaled/2-1+0.05, label=monthly_precip %>% round), vjust=0.05) +
  geom_ma(size = 3, ma_fun = SMA, n = 12, data = scaled, aes(x=Date, y=value, colour = index, linetype=index)) +
#  geom_ma(size = 2, ma_fun = SMA, n = 12, data = scaled, aes(x=Date, y= EVI2_fitted*0.8, colour = "EVI2", linetype="dashed")) +
  theme_light() +
  scale_color_manual(name = "Legend", values= colors) +
  scale_linetype_manual(name = "Legend", values= linetypes) +
  scale_y_continuous(~.*S1scale+S1center) +
  scale_x_date(date_labels = "%b, %Y"
               , limits = xlims
               , date_breaks = "3 month"
#               , minor_breaks = "1 day"
               , expand = expansion(add = 0)) +
  scale_y_continuous(breaks = pretty_breaks()
              , name = "Sentinel-1 RVI Index (detrended)" 
              , sec.axis = sec_axis(~.*S2scale+S2center,name="Sentinel-1 EVI2 Index (detrended)"
              , breaks = pretty_breaks())) +
  xlab("Month, Year") +
  theme(legend.position = "top", legend.title = element_blank(), legend.spacing.x = unit(1, 'cm')
             , axis.text.x = element_text(angle = 45, color = "black") 
             , panel.grid.major = element_blank()
             , panel.grid.minor = element_blank()
             , axis.text.y = element_text(color = "white")
             , axis.title.y = element_text(color = "white", margin = margin(r=-15))
             , axis.text.y.right = element_text(color = "white")
             , axis.title.y.right = element_text(color = "white", margin = margin(l=-15))
             , axis.line.y.left=element_line(color = "red")
             , axis.line.y.right=element_line(color = "red")
             , panel.border = element_blank()) +
  geom_vline(color = "red", xintercept = jan1) +
  # geom_text(aes(y = 1, x=ymd(jan1[[1]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
  # geom_text(aes(y = 1, x=ymd(jan1[[2]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
  # geom_text(aes(y = 1, x=ymd(jan1[[3]]), label = paste0("Jan. 1, ", data$Date %>% year %>% unique)), color="red") +
  ggtitle(paste0("Sentinel-1 & Sentinel-2 modeled seasonality over \n", studyarea, " (", hnumber, " harmonic)"))
```

Save plot
```{r}
ggsave(paste0(site, "_fitted_seasonality_scaled_", S1_band, "_", S2_band,"_rainfall_un", hnumber, "_harmonic.png"), units="in", width=10, height=5, dpi=300, device = 'png')
``` 

Plotting: Check fit with raw data ; model comparison (x= time, y = scatter, single & multiple models)
```{r}
# Read original and fitted time series data files
season_directory <- "/timesat/output/third_run/"
S1_dir <- "/raw/s1/1ha/harmonic/"
S2_dir <- "/raw/s2/1ha/harmonic/"
full_sitelist <- c("mata_escura", "monte_pascoal", "rio_doce", "sooretama")
S1_orb82_start_date <- as.Date("2016-01-10")
S1_orb155_start_date <- as.Date("2016-01-03")
S2_start_date <- as.Date("2018-12-17")

S1_fitted_single <- dir(paste0(wd, S1_dir), full.names = T, pattern = paste0(S1_band,"_2016_2022_harmonic.csv")) %>% map(read.csv, row.names = NULL)
S2_fitted_single <- dir(paste0(wd, S2_dir), full.names = T, pattern = paste0(S2_band,"_2016_2022_harmonic.csv")) %>% map(read.csv, row.names = NULL)
S1_fitted_multiple <- dir(paste0(wd, season_directory, "S1"), full.names = T, pattern = paste0(S1_band,"_STL_season.txt")) %>% map(fread, sep=" ")
S2_fitted_multiple <- dir(paste0(wd, season_directory, "S2"), full.names = T, pattern = paste0(S2_band,"_STL_season.txt")) %>% map(fread, sep=" ")

S1fitted_single <- vector(mode = "list", length = 0) ; S2fitted_single <- vector(mode = "list", length = 0)
S1fitted_multiple <- vector(mode = "list", length = 0) ; S2fitted_multiple <- vector(mode = "list", length = 0)
for (site in site.list) {
      #S1 single
      S1fitted_single[[site]] <- S1_fitted_single[[match(site,full_sitelist)]]
      S1fitted_single[[site]] <- S1fitted_single[[site]] %>% mutate(Date = date_parse(system.time_start, format = "%b %d, %Y")) %>% relocate(Date)
      S1fitted_single[[site]] <- S1fitted_single[[site]] %>% select(-system.time_start) %>%
            group_by(Date) %>% summarize(across(1:2, mean, na.rm=T, .names = "{.col}")) %>% ungroup()
      #S1 multiple
      S1fitted_multiple[[site]] <- S1_fitted_multiple[[match(site,full_sitelist)]]
      S1fitted_multiple[[site]] <- S1fitted_multiple[[site]] %>% t() %>% data.frame()
      colnames(S1fitted_multiple[[site]]) <- "fitted_multiple"
      len <- S1fitted_multiple[[site]][,1] %>% length()
      if(site == "rio_doce") { 
        dates <- seq(S1_orb155_start_date, by = "12 days", length.out=len)
      } else { 
        dates <- seq(S1_orb82_start_date, by = "12 days", length.out=len)
      }
      S1fitted_multiple[[site]] <- S1fitted_multiple[[site]] %>% mutate(Date = dates) %>% relocate(Date)
      
      #S2 single
      S2fitted_single[[site]] <- S2_fitted_single[[match(site,full_sitelist)]]
      S2fitted_single[[site]] <- S2fitted_single[[site]] %>% mutate(Date = date_parse(system.time_start, format = "%b %d, %Y")) %>% relocate(Date)
      S2fitted_single[[site]] <- S2fitted_single[[site]] %>% select(-system.time_start) %>%
            group_by(Date) %>% summarize(across(1:2, mean, na.rm=T, .names = "{.col}")) %>% ungroup()
      
      #S2 multiple
      S2fitted_multiple[[site]] <- S2_fitted_multiple[[match(site,full_sitelist)]]
      S2fitted_multiple[[site]] <- S2fitted_multiple[[site]] %>% t() %>% data.frame()
      colnames(S2fitted_multiple[[site]]) <- "fitted_multiple"
      len <- S2fitted_multiple[[site]][,1] %>% length()
      dates <- seq(S2_start_date, by = "5 days", length.out=len)
      S2fitted_multiple[[site]] <- S2fitted_multiple[[site]] %>% mutate(Date = dates) %>% relocate(Date)
}

#Construct multiple time series dataframe for each site 
S1series <- vector(mode = "list", length = 0)
S2series <- vector(mode = "list", length = 0)
for (site in site.list) {
  #S1
  S1series[[site]] <- merge.zoo(read.zoo(S1fitted_single[[site]]), 
                                read.zoo(S1fitted_multiple[[site]]))
  names(S1series[[site]]) <- c(names(S1fitted_single[[site]][-1]), "fitted_multiple")
  #S2
  S2series[[site]] <- merge.zoo(read.zoo(S2fitted_single[[site]]), 
                                read.zoo(S2fitted_multiple[[site]]))
  names(S2series[[site]]) <- c(names(S2fitted_single[[site]][-1]), "fitted_multiple")
}

#define outlier function
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.1, .9), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}


# choose site to plot
site <- "rio_doce"



#par(mfrow = c(2, 1))
S1series[[site]] %>% na.approx() %>% plot(type = c("p", "l", "l"), screens = 1, col = c("blue", "orange", "purple"), cex = 0.6, pch=20, ylab="", xlab="", cex.axis = 1.5, main = "")
mtext(expression(bold("Sentinel-1 RVI")), side = 2, line = 2.2, col = "blue" , adj=0.5, cex = 1.7)
mtext("Date", side = 1, line = 4, cex = 2)

S2series[[site]] %>% na.approx() %>% plot(type = c("p", "l", "l"), screens = 1, col = c("darkgreen", "orange", "purple"), cex = 0.6, pch=20, ylim = c(0.2,0.6), ylab="", xlab="", cex.axis = 1.5, main = "")
mtext(expression(bold("Sentinel-2 EVI2")), side = 2, line = 2.2, col = "darkgreen", adj=0.5, cex = 1.7)
mtext("Date", side = 1, line = 4, cex = 2)

```

A/B. Plot setup - annual fitted time series S1/S2 (normal (Jan.-Dec.) and recentered (April-March))
```{r}
#Define study site
study_site <-"sooretama" # "sooretama" # "rio_doce" # "monte_pascoal" # "mata_escura"

if (study_site == "rio_doce") {studyarea <-  "PE Rio Doce, MG"
} else if (study_site == "sooretama") {studyarea <- "REBIO Sooretama, ES"
} else if (study_site == "monte_pascoal") {studyarea <- "PNH Monte Pascoal, BA"
} else if (study_site == "mata_escura") {studyarea <- "REBIO Mata Escura, MG"
} else {studyarea <- "NA"}

if (study_site == "rio_doce" | study_site == "mata_escura") {foresttype <- "seasonal semi-deciduous forest"
} else if (study_site == "monte_pascoal" | study_site == "sooretama") {foresttype <- "broadleaf evergreen forest"
} else {foresttype <- "NA"}

S1fit <- fittedlist[[study_site]][[1]] %>% data.frame(); S2fit <- fittedlist[[study_site]][[2]] %>% data.frame()
S1tr <- trendlist[[study_site]][[1]] %>% data.frame(); S2tr <- trendlist[[study_site]][[2]] %>% data.frame()

#detrend and scale
rownames(S1fit) <- S1fit$Date ; rownames(S2fit) <- S2fit$Date
rownames(S1tr) <- S1tr$Date ; rownames(S2tr) <- S2tr$Date

#Sentinel-1 convert fit and trend to xts 
S1fit_xts <- S1fit %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S1fit), "%Y-%m-%d"))
S1tr_xts <- S1tr %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S1tr), "%Y-%m-%d"))

#Sentinel-2 convert fit and trend to xts 
S2fit_xts <- S2fit %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S2fit), "%Y-%m-%d"))
S2tr_xts <- S2tr %>% select(-c(Date)) %>% xts(order.by = as.Date(rownames(S2tr), "%Y-%m-%d"))
S1detr_xts <- S1fit_xts - S1tr_xts ; S2detr_xts <- S2fit_xts - S2tr_xts

S1scaled <- S1detr_xts %>% scale()
S1scale <- scale(S1detr_xts) %>% attr("scaled:scale") ; S1center <- scale(S1detr_xts)  %>% attr("scaled:center")
S2scaled <- S2detr_xts %>% scale()
S2scale <- scale(S2detr_xts) %>% attr("scaled:scale") ; S2center <- scale(S2detr_xts)  %>% attr("scaled:center")
scaled <- merge(S1scaled, S2scaled) 
#  %>% pivot_longer(!Date, names_to = "series") %>% mutate(Date = as.Date(Date))

#data <- data.frame(dates, S1fit, S2fit) 

S1legend <- paste0("all plots mean", S1_band) ; S2legend <- paste0("all plots mean", S2_band)
colors <- c("all plots mean annual RVI" = "blue", "all plots mean annual EVI2" = "dark green")
linetypes <- c("all plots mean annual RVI" = "solid", "all plots mean annual EVI2" = "solid")   # doesn't seem to be working

xlims <- c(as.Date("2019-01-01"), as.Date("2022-01-01"))
#ylims <- c(-10,10)

scaled <- fortify(scaled)
colnames(scaled)[] <- c("Date", paste0(S1_band, "_fitted"), paste0(S2_band, "_fitted"))
jan1 <- paste0(scaled$Date %>% year %>% unique, "-01-01") %>% as.Date()

scaled_means <- scaled  # %>% filter(between(Date, xlims[1], xlims[2])) %>% mutate(year = year(Date))
scaled_means <- scaled_means %>% mutate(doy = ymd(paste0("1901-", month(Date), "-", day(Date)))) %>% group_by(doy) %>% summarise_all(.funs = c(mean="mean"), na.rm = TRUE) %>% pad(by='doy') 
scaled_means$RVI_fitted_mean[is.nan(scaled_means$RVI_fitted_mean)]<-NA
scaled_means$EVI2_fitted_mean[is.nan(scaled_means$EVI2_fitted_mean)]<-NA

# reorder table to center the peak in the plot
scaled_means_high <- scaled_means %>% filter(doy >= as.Date("1901-04-01"))  # April 1st cutoff
scaled_means_low <- scaled_means %>% filter(doy < as.Date("1901-04-01"))
scaled_means_reorder <- rbind(scaled_means_high, scaled_means_low) %>% select(-Date_mean)

#import ggplot themes (for "ipsum" theme)
hrbrthemes::update_geom_font_defaults()

#create x-axis labels
months <- as.Date(scaled_means_reorder$doy, origin = "1901-01-01") %>% month()
labels <- month.name[months] %>% substr(1,3) %>% paste0("-01") %>% data.frame(cbind(1:length(months)), row.names = NULL) ; colnames(labels) <- c("label", "id")
numbers <- labels %>% group_by(label) %>% summarise(number = n(), sum = sum(id)) %>% arrange(sum)
labels <- numbers %>% rowwise() %>% mutate(nice_labels = list(c(label, rep(NA, len=number-1)))) %>% unnest_longer(nice_labels) %>% select(nice_labels)

scaled_means_reorder <- cbind(scaled_means_reorder,labels) %>% rownames_to_column() %>% mutate(nice_labels = replace_na(nice_labels, " "))
scaled_means_reorder$rowname <- factor(scaled_means_reorder$rowname, levels = scaled_means_reorder$rowname, ordered = T)

scaled_means_reorder["group"] <- rep(1,length(scaled_means_reorder$rowname))
scaled_means_reorder <- scaled_means_reorder %>% select(-doy)
```

A/B. Plot mean annual fitted time series S1/S2 (normal)
```{r}
# plot mean annual
data <- scaled %>% mutate(doy = ymd(paste0("1901-", month(Date), "-", day(Date))))
ggplot() +  
  geom_ma(ma_fun = ZLEMA, n = 20, data= scaled_means[!is.na(scaled_means$RVI_fitted_mean),]
            , aes(group =1, x=doy, y= RVI_fitted_mean, colour = "all plots mean annual RVI", linetype = "solid"), size = 0.8) +
 # geom_ma(ma_fun = ZLEMA, n = 20, data = scaled_means[!is.na(scaled_means$EVI2_fitted_mean),]
 #           , aes(group =1, x=doy, y= EVI2_fitted_mean, colour = "all plots mean annual EVI2", linetype = "solid"), size = 0.8) +
  #geom_line(data= scaled_means[!is.na(scaled_means$RVI_fitted_mean),]   # strange that the line geom gives correct linetype
  #          , aes(group =1, x=doy, y= RVI_fitted_mean, colour = "all plots mean annual RVI"), size = 0.8) +
  #geom_line(data = scaled_means[!is.na(scaled_means$EVI2_fitted_mean),]
  #          , aes(group =1, x=doy, y= EVI2_fitted_mean, colour = "all plots mean annual EVI2"), size = 0.8) +
  #geom_point(data = data, aes(x=doy, y= RVI_fitted, colour = "all plots mean annual RVI"), size = 0.3, alpha = 0.4) +
  #geom_point(data = data, aes(x=doy, y= EVI2_fitted, colour = "all plots mean annual EVI2"), size = 0.3, alpha = 0.4) +
  #geom_path(aes(x=rowname, y= EVI2_fitted_mean, colour = "all plots mean EVI2"), size = 0.8) +
  geom_vline(color = "black", alpha = 0.1, xintercept = jan1) +
  #geom_line(data= scaled_means[!is.na(scaled_means$RVI_fitted_mean),]
  #          , aes(group =1, x=rowname, y= RVI_fitted_mean, colour = "all plots mean RVI"), size = 0.8) +
  #geom_line(data = scaled_means[!is.na(scaled_means$EVI2_fitted_mean),]
  #          , aes(group =1, x=rowname, y= EVI2_fitted_mean, colour = "all plots mean EVI2"), size = 0.8) +
  scale_color_manual(name = " ", values= colors) +
  #scale_linetype_manual(values= linetypes) +
  theme_bw() +
  scale_x_date(name = "Day of Year", date_labels = "%b-%d", date_breaks = "1 month", expand = expansion(add = .05)) +
  scale_y_continuous(~.*S1scale+S1center, breaks = pretty_breaks()
              , name = "Sentinel-1 RVI Index (detrended)" 
              , sec.axis = sec_axis(~.*S2scale+S2center, name="Sentinel-2 EVI2 Index (detrended)"
              , breaks = pretty_breaks())) +
  theme(legend.position = "top", 
         panel.grid.major.x = element_blank(),
         axis.text.y = element_text(color = "blue"),
         axis.text.x = element_text(vjust = 0.5, angle = 60),
         axis.title.y = element_text(color = "blue"),
         axis.text.y.right = element_text(color = "dark green"),
         axis.title.y.right = element_text(color = "dark green"),
         panel.grid.major.y =  element_blank(),
         panel.grid.minor = element_blank()) +
   ggtitle(paste0("Annual means (", year(xlims)[1], "-", year(xlims)[2], ") of modeled seasonality (", hnumber, " harmonic)\nover ", foresttype, ", ", studyarea, " "))
```

Save normal plot
```{r}
ggsave(paste0(study_site, "_mean_annual_seasonality_", S1_band, "_", S2_band,"_", hnumber, "_harmonic.png"), units="in", width=10, height=5, dpi=300, device = 'png')
```

A/B. Plot annual fitted time series S1/S2 (recentered)
```{r}
# plot mean annual, reordered
ggplot() +  
  geom_ma(ma_fun = ZLEMA, n = 20, data= scaled_means_reorder[!is.na(scaled_means_reorder$RVI_fitted_mean),]
            , aes(group =1, x=rowname, y= RVI_fitted_mean, colour = "all plots mean annual RVI"), size = 0.8) +
  geom_ma(ma_fun = ZLEMA, n = 20, data = scaled_means_reorder[!is.na(scaled_means_reorder$EVI2_fitted_mean),]
            , aes(group =1, x=rowname, y= EVI2_fitted_mean, colour = "all plots mean annual EVI2"), size = 0.8) +
  #geom_path(aes(x=rowname, y= EVI2_fitted_mean, colour = "all plots mean EVI2"), size = 0.8) +
  geom_vline(color = "black", alpha = 0.1, xintercept = c(0, cumsum(numbers$number))) +
  #geom_line(data= scaled_means_reorder[!is.na(scaled_means_reorder$RVI_fitted_mean),]   # strange that the line geom gives correct linetype
  #          , aes(group =1, x=rowname, y= RVI_fitted_mean, colour = "all plots mean RVI"), size = 0.8) +
  #geom_line(data = scaled_means_reorder[!is.na(scaled_means_reorder$EVI2_fitted_mean),]
  #          , aes(group =1, x=rowname, y= EVI2_fitted_mean, colour = "all plots mean EVI2"), size = 0.8) +
  scale_color_manual(name = " ", values= colors) +
  #scale_linetype_manual(values= linetypes) +
  theme_bw() +
  scale_x_discrete(labels = scaled_means_reorder$nice_labels
                   , name = "Day of Year"
                 #  , breaks = c(0, cumsum(numbers$number))  # fix with secondary axis???
                   ) +
  scale_y_continuous(~.*S1scale+S1center, breaks = pretty_breaks()
              , name = "Sentinel-1 RVI Index (detrended)" 
              , sec.axis = sec_axis(~.*S2scale+S2center, name="Sentinel-2 EVI2 Index (detrended)"
              , breaks = pretty_breaks())) +
  theme(legend.position = "top", 
         panel.grid.major.x = element_blank(),
         axis.text.y = element_text(color = "blue"),
         axis.text.x = element_text(vjust = 0.5, angle = 60),
         axis.title.y = element_text(color = "blue"),
         axis.text.y.right = element_text(color = "dark green"),
         axis.title.y.right = element_text(color = "dark green"),
         panel.grid.major.y =  element_blank(),
         panel.grid.minor = element_blank()) +
   ggtitle(paste0("Annual means (", year(xlims)[1], "-", year(xlims)[2], ") of modeled seasonality (", hnumber, " harmonic)\nover ", foresttype, ", ", studyarea, " "))
```

Save recentered plot
```{r}
ggsave(paste0(study_site, "_mean_annual_seasonality_", S1_band, "_", S2_band,"_", hnumber, "_harmonic_recentr.png"), units="in", width=10, height=5, dpi=300, device = 'png')
```

Other plots...

(plot entire time series - fits)
```{r}
ggplot(data = scaled, aes(x=Date)) +  
  geom_ma(ma_fun = SMA, n = 20, aes(y= RVI_fitted, colour = "all plots mean RVI")) +
  geom_ma(ma_fun = SMA, n = 20, aes(y= EVI2_fitted, colour = "all plots mean EVI2")) +
  #geom_line(aes(y= RVI_fitted, colour = "all plots mean RVI")) +
  #geom_line(aes(y= EVI2_fitted, colour = "all plots mean EVI2")) +
  geom_vline(color = "red", alpha = 0.3, xintercept = jan1) +
  scale_color_manual(name = "Legend", values= colors) +
  #scale_linetype_manual(values= linetypes) +
  theme_bw() +
  scale_x_date(date_labels = "%b,%Y", limits = xlims, date_breaks = "3 month", expand = expansion(add = .05)) +
  scale_y_continuous(~.*S1scale+S1center, breaks = pretty_breaks()
              , name = "Sentinel-1 RVI Index (detrended)"
              , sec.axis = sec_axis(~.*S2scale+S2center, name="Sentinel-1 EVI2 Index (detrended)"
              , breaks = pretty_breaks())) +
  xlab("Month, Year") +
  theme(legend.position = "top", 
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.text.y = element_text(color = "blue"),
  axis.text.x = element_text(vjust = 0.5, angle = 60),
  axis.title.y = element_text(color = "blue"),
  axis.text.y.right = element_text(color = "dark green"),
  axis.title.y.right = element_text(color = "dark green")) +
  ggtitle(paste0("Sentinel-1 & Sentinel-2 modeled seasonality over \n", site, " (multiple harmonic)"))
```

Split data using season definition defined above (A or B)
```{r}
plotdata.stats <- plotdata %>% 
              mutate(day = yday(date)) %>%  #create a month factor
              mutate(season=cut(day, #create season factor
              breaks=season_def,  #define seasons by day of year 
              labels=c("wet","dry","wet"))) %>%  
              mutate(year = year(date)) %>%  #create a year factor
              mutate(Hectare_ID = substring(Hectare_ID, 14)) %>%
              group_by(Hectare_ID) %>%    #group and reorder ascending
              mutate(mean = mean(value, na.rm = TRUE)) %>% arrange(mean) %>%
             # mutate(sortorder = rep(seq(1:length(IDs)), length.out = length(Hectare_ID))) %>%
              ungroup()

plotdata.stats <- data.frame(value = plotdata.stats$value,      #rearrange
                             Hectare_ID = factor(plotdata.stats$Hectare_ID),
                             season = factor(plotdata.stats$season), 
                             year = factor(plotdata.stats$year)
#                             ,sortorder = factor(indexdata.stats$sortorder)
)
```

Plot seasonality by 1-ha plot
```{r}
#take a random sample 
selectedrows <- sample.int(plotdata.stats$Hectare_ID %>% unique() %>% length(), 30, replace = F)
plotdata.stats <- plotdata.stats[is.element(plotdata.stats$Hectare_ID, selectedrows),]

#plot seasonal comparison
xlabs <- plotdata.stats$Hectare_ID %>% unique() %>% as.character()
ylims <- c(-16,-10)
#check boxplots
colors = c("wet" = "blue", "dry" = "orange")
plotdata.stats %>% na.omit() %>% 
  ggplot(aes(y=value, x=Hectare_ID, color = season, fill = season)) + 
    geom_boxplot(outlier.size=0.5, outlier.colour = NULL, show.legend = FALSE) + 
    scale_fill_manual(values=c("blue", "orange")) +
    scale_color_manual(values=c("blue", "orange")) +
    scale_x_discrete(labels=xlabs) +
    ylab(paste0("Mean ", bandname, " index value (dB)")) +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 90, vjust=0.5)) +
    ggtitle(paste0("Plot-wise seasonal comparison of mean annual ", bandname, " backscatter \n (Sentinel-1) for ",  
                   "30 randomly selected 1 ha",
    #     length(unique(plotdata$Hectare_ID)), " ", scale, 
         " plots, ", studyarea, ", Brazil"))
```

Save plot
```{r}
ggsave(paste0(studyarea_base, "_plotseasonal_s1_", bandname, ".png"), units="in", width=10, height=5, dpi=300, device = 'png')
```

Analysis of variance:season
```{r}
#fit the linear model 
fit <- lm(value ~ season*year*Hectare_ID, plotdata.stats, 
          contrasts = list(Hectare_ID = "contr.sum", year = "contr.poly"))
anova(fit) %>% kable(digits = 10) 
write.csv(anova(fit) , paste0(studyarea_base, "_anova_s1_", bandname, ".csv"))

```

Histogram
```{r}
#plot histogram
hist(indexdata.mean$mean, breaks = 20, ylim = c(0,25), xlab= paste0("Mean ", bandname, " value (dB)"), ylab= "Number of days / year", main= paste0("\n\n Distribution of daily Sentinel-1 averaged ", bandname, " backscatter \n values for ", length(indexdata)-1, " 1-ha plots over ", foresttype, ",\n ", studyarea, ", Brazil (2016-2021)"), border=F, col="dark blue")

abline(h=seq(0,55,5), col="gray", lty="dotted")

ggsave(paste0(studyarea_base, "_histannual_s1.png"), units="in", width=10, height=5, dpi=300, device = 'png')
```

misc code
```{r}
if (study_site == "rio_doce") {studyarea <-  "PE Rio Doce, MG"
} else if (study_site == "sooretama") {studyarea <- "REBIO Sooretama, ES"
} else if (study_site == "monte_pascoal") {studyarea <- "PNH Monte Pascoal, BA"
} else if (study_site == "mata_escura") {studyarea <- "REBIO Mata Escura, MG"
} else {studyarea <- "NA"}

if (study_site == "rio_doce" | study_site == "mata_escura") {foresttype <- "seasonal semi-deciduous forest"
} else if (study_site == "monte_pascoal" | study_site == "sooretama") {foresttype <- "broadleaf evergreen forest"
} else {foresttype <- "NA"}
```

